%
% Chapter 2
%
\chapter {Basics of Quantum Computing}

\section{Dirac Notation}

Also known as bra-ket notation, Dirac notation provides a convenient way of expressing the vectors used in quantum mechanics.

Dirac notation defines two elements:

\begin{itemize}
    \item The \textit{ket} ($\ket{\psi}$), which denotes a column vector in a complex vector space that represents a quantum state.
    $$\ket{\psi}=\left[\begin{array}{c}\psi_{1} \\ \psi_{2} \\ \cdots \\ \psi_{n} \end{array}\right]$$
    \item The \textit{bra} ($\bra{\psi}$), which denotes a row vector that is the conjugate transpose, or adjoint, of a corresponding \textit{ket} ($\ket{v}$).
    $$\bra{\psi}=\left[\begin{array}{cccc}\psi_{1}^{*} & \psi_{2}^{*} & \cdots & \psi_{n}^{*}\end{array}\right]=\left[\begin{array}{c}\psi_{1} \\ \psi_{2} \\ \cdots \\ \psi_{n} \end{array}\right]^{\dagger}$$
\end{itemize}

An \textit{operator} $\bf{A}$, represented by a $n \times n$ matrix, acting on a \textit{ket} $\ket{\psi}$ produces another \textit{ket} $\ket{\psi'}$ such that the produced \textit{ket} can be computed by matrix multiplication:
$$\ket{\psi'}=\textbf{A}\ket{\psi}=\left[\begin{array}{cc}A_{11} & A_{12} \\ A_{21} & A_{22}\end{array}\right]\left[\begin{array}{c}\psi_{1} \\ \psi_{2}\end{array}\right]=\left[\begin{array}{c}A_{11}\psi_{1}+A_{12}\psi_{2} \\ A_{21}\psi_{1}+A_{22}\psi_{2}\end{array}\right]$$

The \textit{inner product} is denoted as a bra-ket pair $\braket{\phi}{\psi}$ and represents the probability amplitude that a quantum state $\psi$ would be subsequently found in state $\phi$:
$$\braket{\phi}{\psi}=\left[\begin{array}{cc}\phi_{1}^{*} & \phi_{2}^{*} \end{array}\right]\left[\begin{array}{c}\psi_{1} \\ \psi_{2} \end{array}\right]=\phi_{1}^{*}\psi_{1}+\phi_{2}^{*}\psi_{2}$$

This notation also provides a way to describe the state vector of $n$ uncorrelated quantum states, the \textit{tensor product}:
$$\ket{\phi}\otimes\ket{\psi}=\ket{\phi}\ket{\psi}=\ket{\phi\psi}=\left[\begin{array}{c}\phi_{1} \\ \phi_{2}\end{array}\right]\otimes\left[\begin{array}{c}\psi_{1} \\ \psi_{2}\end{array}\right]=\left[\begin{array}{c}\phi_{1}\psi_{1} \\ \phi_{1}\psi_{2} \\ \phi_{2}\psi_{1} \\ \phi_{2}\psi_{2}\end{array}\right]$$

Note that $\ket{\psi}^{\otimes n}$ represents the tensor product of $n$ $\ket{\psi}$ quantum states:
$$\ket{\psi}^{\otimes n}=\ket{\psi}\otimes\cdots\otimes\ket{\psi}=\ket{\psi}\cdots\ket{\psi}=\ket{\psi\cdots\psi}=\left[\begin{array}{c}\psi_{1} \\ \psi_{2} \end{array}\right]\otimes\cdots\otimes\left[\begin{array}{c}\psi_{1} \\ \psi_{2} \end{array}\right]$$

\section{Quantum Systems}

Quantum mechanics is a mathematical framework for the development of physical theories\cite{NielsenChuang_2012}. On its own quantum mechanics doesn't tell you what laws a physical system must obey, but it does provide a mathematical and conceptual framework for the development of such laws. The postulates described next provide a connection between the physical world and the mathematical formalism of quantum mechanics.

Postulate 1: The state of an isolated physical system is fully described by a unit vector $\ket{\psi}$ in Hilbert space.

Postulate 2: The evolution of a closed quantum system is described by a unitary transformation. This means that the state $\ket{\psi}$ at time $t_1$ is related to the state $\ket{\psi'}$ at time $t_2$ by a unitary operator $U$ which depends only on the times $t_1$ and $t_2$:

$$\ket{\psi'}=U(t_1,t_2)\ket{\psi}$$

Postulate 3: The state space of a composite physical system is the tensor product of the state spaces of the component physical systems.

Postulate 4: Quantum measurements are described by a collection of ${M_m}$ of measurement operators.

\section{Qubits}

In classical information theory and classical computing, a bit is the fundamental building block.
Analogously, in quantum information theory and quantum computing, a quantum bit or qubit is the fundamental building block.

Physically, a quibit is a two-level quantum-mechanical system. The polarization of a single photon and the spin of the electron are examples of such systems.

Mathematically, a qubit is a linear combination of states $\ket{\psi}=\alpha\ket{0}+\beta\ket{1}$ where $\alpha$ and $\beta$ are complex numbers known as amplitudes, and $\ket{0}$ and $\ket{1}$ are the computational basis states.

When a qubit is measured, the result is either $\ket{0}$ with probability $|\alpha|^2$ or $\ket{1}$ with probability $|\beta|^2$.
Since the probabilities must sum to one, the qubit's state is normalized:
$$|\alpha|^2+|\beta|^2=1$$

The computational basis states, which are analogous to the two values ($0$ and $1$) that a classical bit may take, form an orthonormal basis represented by the following vectors:
$$\ket{0}=\left[\begin{array}{c}1 \\ 0\end{array}\right]$$
$$\ket{1}=\left[\begin{array}{c}0 \\ 1\end{array}\right]$$

Using the previous definitions, we can see a qubit as a unit vector in two-dimensional complex vector space:
\[\ket{\psi}=\left[\begin{array}{c}\alpha \\ \beta\end{array}\right]\]

\section{Superposition}

The principle of quantum superposition states that the most general state of a quantum-mechanical system is a linear combination of all distinct valid quantum states. A qubit is an example of a quantum superposition of the basis states $\ket{0}$ and $\ket{1}$.

A concrete example of a qubit in superposition that has the same probability of being measured as $\ket{0}$ or $\ket{1}$ is the following:
$$\ket{\psi}=\frac{1}{\sqrt{2}}\left(\ket{0}+\ket{1}\right)$$

\section{Quantum Logic Gates}

In classical digital circuits, logic gates are the building blocks. Analogously, in the quantum circuit model of computation, quantum logic gates are the building blocks of quantum algorithms. Quantum gates act on qubits, transform them in different ways, and can be applied sequentially to perform complex quantum computations.

Quantum gates are unitary operators represented as $2^n \times 2^n$ unitary matrices where $n$ is the number of qubits the gate operates on. Unitary matrices are complex square matrices $\bf{U}$ which have the property that its conjugate transpose or adjoint $\bf{U^{\dagger}}$ is also its inverse $\bf{U^{-1}}$:
$$\bf{U^{\dagger}U}=\bf{UU^{\dagger}}=\bf{UU^{-1}}=I$$

A quantum gate is applied to a qubit system by multiplying the gate's matrix representation by the qubits' state vector. This operation transforms the qubit system:
$$\ket{\psi_1}=\bf{U}\ket{\psi_0}$$

Applying a sequence of quantum gates is equivalent to performing a series of these multiplications. For example, applying gate $\bf{U_a}$ followed by a gate $\bf{U_b}$ to a state vector $\ket{\psi}$ is represented by the follwing expression where the gates closest to the state vector are applied first:
$$\bf{U_b}\bf{U_a}\ket{\psi}$$

Since matrix multiplication is associative, multiplying $\bf{U_a}$ by $\bf{U_b}$ produces a compund gate $\bf{U_{b}U_{a}}$ that is equivalent to applying $\bf{U_a}$ followed by $\bf{U_b}$:
$$\bf{U_b}\bf{U_a}\ket{\psi}=\bf{U_b}(\bf{U_a}\ket{\psi})=(\bf{U_b}\bf{U_a})\ket{\psi}$$

Note that all quantum gates are reversible since they are represented by unitary matrices. This means that for any gate, another gate exists that reverts the gate's transformation on a state vector:
$$\ket{\psi}=\bf{U^{\dagger}}(\bf{U}\ket{\psi})=\bf{U^{\dagger}}\bf{U}\ket{\psi}=\bf{I}\ket{\psi}$$

\section{Single-Qubit Gates}

Single-qubit gates are represented by $2 \times 2$ unitary matrices. Applying a gate to a matrix can be done by multiplying the gate's matrix by the qubit's state vector.

Next, some of the most commonly used gates.

Pauli-X gate:
$$X=\left[\begin{array}{cc}0 & 1 \\ 1 & 0\end{array}\right]$$

Pauli-Y gate:
$$Y=\left[\begin{array}{cc}0 & -\imath \\ \imath & 0\end{array}\right]$$

Pauli-Z gate:
$$Z=\left[\begin{array}{cc}1 & 0 \\ 0 & -1\end{array}\right]$$

Hadamard gate:
$$H=\frac{1}{\sqrt{2}}\left[\begin{array}{cc}1 & 1 \\ 1 & -1\end{array}\right]$$

RX gate:
$$R_x(\theta)=\left[\begin{array}{cc}cos{\frac{\theta}{2}} & -\imath sin{\frac{\theta}{2}} \\ -\imath sin{\frac{\theta}{2}} & cos{\frac{\theta}{2}}\end{array}\right]$$

RY gate:
$$R_y(\theta)=\left[\begin{array}{cc}cos{\frac{\theta}{2}} & -sin{\frac{\theta}{2}} \\ sin{\frac{\theta}{2}} & cos{\frac{\theta}{2}}\end{array}\right]$$

RZ gate:
$$R_z(\theta)=\left[\begin{array}{cc}e^{-\imath \frac{\theta}{2}} & 0 \\ 0 & e^{\imath \frac{\theta}{2}}\end{array}\right]$$

R1 gate:
$$R_1(\theta)=\left[\begin{array}{cc}1 & 0 \\ 0 & e^{\imath \theta}\end{array}\right]$$

\section{Multi-Qubit Gates}

Multi-qubit gates are represented by $2^N \times 2^N$ matrices where N is the number of qubits the gate operates on.

Of special importance is the CNOT gate which is a two-qubit gate where the first qubit is referred to as the control qubit and the second qubit as the target qubit. CNOT acts as a conditional gate, if the control qubit is in state $\ket{1}$, it applies the Pauli X gate to the target qubit, otherwise it does nothing. The main use of this gate is to prepare entangled states.

$$CNOT=\left[\begin{array}{cccc}1 & 0 & 0 & 0 \\ 0 & 1 & 0 & 0 \\ 0 & 0 & 0 & 1 \\ 0 & 0 & 1 & 0\end{array}\right]$$
