%
% Chapter 5
%
\chapter {Analysis Framework}

The strategy we use to analyze Shor's integer factorization algorithm is very similar to the one proposed by Soeken et al.\cite{ResourceEstimationFramework_Soeken_2021}. The overview of the process is the following:
\begin{enumerate}
    \item Implement the quantum algorithm using a high-level programming language.
    \item Verify the correctness of the implementation by executing the algorithm in a full state simulator.
    \item Use, extend and build tools that can use the high-level algorithm implementation to estimate the amount of logical resources the algorithm requires depeding on its input size.
    \begin{enumerate}
        \item As part of logical resources estimation, consider the effects of introducing error-corrected qubits and fault-tolerant quantum gates on the amount of required resources.
    \end{enumerate}
    \item For each physical platform do the following:
    \begin{enumerate}
        \item Create a tool that uses platform specific parameters to calculate the amount of physical qubits, physical gates, and runtime required to run the algorithm.
        \item Analyze the data produced by the tool to determine the following:
        \begin{enumerate}
            \item The maximum input size for the algorithm to run on a real NISQ device.
            \item The characteristics that a device should have to run the algorithm for a specific input size, and determine what would be the runtime.
        \end{enumerate}
    \end{enumerate}
    \item Compare the results obtained for each physical platform.
\end{enumerate}

\section{Q\# and the Quantum Development Kit (QDK)}

Microsoft's Quantum Development Kit (QDK) supports the implementation of custom simulators that can be used to run Q\# programs. We leverage this capability and implement a simulator that calculates the resources a quantum algorithm would require to be executed in a hardware platform with specific characteristics.

\todo{Explain the following}
\begin{itemize}
    \item QDK Custom Simulators: Describe how custom simulators are implemented using diagrams and code snippets.
    \item Software Architecture of Physical Resources Estimator Simulator: Describe the software architecture using diagrams and code snippets.
\end{itemize}

Source code of a working version can be found in in \href{https://github.com/cesarzc/uw-master-in-physics-project}{GitHub}.
